### 2019-11-27 scripts/manuscript.Rmd

---
title             : "Personality and political beliefs across the lifespan"
shorttitle        : "PERSONALITY AND POLITICAL BELIEFS"
author: 
  - name          : "Sarah Dimakis"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "University of Oregon, Eugene, OR"
    email         : "sdimakis@uoregon.edu"
  - name          : "Meghan Siritzky"
    affiliation   : "1"
  - name          : "Jamie Yellowtail"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "University of Oregon"
authornote: |
  This project was completed as part of the EDLD Introduction to Data Science class at the University of Oregon.
abstract: |
  Here is where we will write our abstract.

keywords          : "keywords"
wordcount         : "X"
bibliography      : ["r-references.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library(papaja)
library(rio)
library(here)
library(tidyverse)
library(psych)

data <- import(here("data", "EPB.sav"), setclass = "tbl_df")
```

```{r analysis-preferences}
knitr::opts_chunk$set(echo = FALSE)

```

```{r filter}
#Removing particpants from the sample

#filter out preview mode (not real participants) and those who did not complete the survey
data <- data %>% 
  filter(DistributionChannel != "preview" & Progress == 100)

#filter out those who did not pass the attention checks
data <- data %>% 
  filter(Q6_8 == 6 & Q5_14 == 3)

#filter out those who took the survey too quickly (the time cut offs were predetermined and based on which condition the participant was randomly assigned to)
data <- data %>% 
  pivot_longer(cols = Video_DO_V4:Video_DO_V1, 
               names_to = "Video_DO", 
               values_to = "Value",
               names_prefix = "Video_DO_",
               values_drop_na = TRUE) %>% 
  filter(Video_DO == 'V1' & Duration__in_seconds_ >316 | 
           Video_DO == 'V2' & Duration__in_seconds_ > 340 | 
           Video_DO == 'V3' & Duration__in_seconds_ > 398 |
           Video_DO == 'V4' & Duration__in_seconds_ > 666 | 
           Video_DO == 'V5' & Duration__in_seconds_ > 565 | 
           Video_DO == 'V6' & Duration__in_seconds_ > 538) %>% 
  filter(Duration__in_seconds_ < 5000) #this person spent too long taking the survey
```

```{r aggregrateSEC}

#Aggregating the social and economic conservationism scale (SECs) using scoreItems from the psych package and mutate from dplyr
items <- c("Q3_1", "Q3_2","Q3_3","Q3_4","Q3_5","Q3_6","Q3_7","Q3_8","Q3_9","Q3_10","Q3_11","Q3_12" )

#econonmic conservatism subscale 
scaleKey <- c(0, 1, 0, 0, -1, 1, 0, 0, 1, 1, 0, 0)
economic <- scoreItems(keys = scaleKey, items = data[items], 
  totals = FALSE, missing = FALSE, min = 0, max = 100)
data <- data %>% 
  mutate(SEC_economic_conservatism = economic$scores[,1])

#social conservatism subscale
scaleKey <- c(-1, 0, 1, 1, 0, 0, 1, 1, 0, 0, 1, 1)
social <- scoreItems(keys = scaleKey, items = data[items], 
  totals = FALSE, missing = FALSE, min = 0, max = 100)
data <- data %>% 
  mutate(SEC_social_conservatism = social$scores[,1])
```

```{r aggregateTIPI}

#Aggregating big 5 traits (TIPI) using scoreItems from the psych package and mutate from dplyr
items <- c("Q4_1", "Q4_2", "Q4_3", "Q4_4", "Q4_5", "Q4_6", "Q4_7", "Q4_8", "Q4_9", "Q4_10")

#Extraversion
scaleKey <- c(1, 0, 0, 0, 0, -1, 0, 0, 0, 0)
extraversion <- scoreItems(keys = scaleKey, items = data[items], 
  totals = FALSE, missing = FALSE, min = 1, max = 7)
data <- data %>% 
  mutate(Extraversion = extraversion$scores[,1])

#Agreeableness
scaleKey <- c(0, -1, 0, 0, 0, 0, 1, 0, 0, 0)
agree <- scoreItems(keys = scaleKey, items = data[items], 
  totals = FALSE, missing = FALSE, min = 1, max = 7)
data <- data %>% 
  mutate(Agreeableness = agree$scores[,1])

#Conscientiousness
scaleKey <- c(0, 0, 1, 0, 0, 0, 0, -1, 0, 0)
consc <- scoreItems(keys = scaleKey, items = data[items], 
  totals = FALSE, missing = FALSE, min = 1, max = 7)
data <- data %>% mutate(Conscientiousness = consc$scores[,1])

#Emotional stability
scaleKey <- c(0, 0, 0, -1, 0, 0, 0, 0, 1, 0)
emo <- scoreItems(keys = scaleKey, items = data[items], 
  totals = FALSE, missing = FALSE, min = 1, max = 7)
data <- data %>% 
  mutate(Emotional_stability = emo$scores[,1])

#Openness to experiences
scaleKey <- c(0, 0, 0, 0, 5, 0, 0, 0, 0, -1)
open <- scoreItems(keys = scaleKey, items = data[items], 
  totals = FALSE, missing = FALSE, min = 1, max = 7)
data <- data %>% 
  mutate(Openness_to_experiences = open$scores[,1])

```

```{r select}
#selecting the variables that we want to use, cleaning, and characterizing dataset
data <- data %>% 
  select(ResponseId, Q7, Q8, Q9, Q10, Q11,
         SEC_economic_conservatism, SEC_social_conservatism,
         Agreeableness, Conscientiousness, Openness_to_experiences, Extraversion, Emotional_stability) %>% 
  rename(Gender = Q7, 
         Age = Q8, 
         Ethnicity = Q9, 
         Political_orientation = Q10,
         Political_affiliation = Q11,
         id = ResponseId) %>% 
  janitor::clean_names() %>% 
  characterize() %>% 
  mutate(age = as.numeric(age))

#Are you going to include the ethnicity data in later plots? If not, you could just select the other columns. 
```

```{r tidying}
#last step is to tidy our data
tidy <- data %>% 
  pivot_longer(cols = c(sec_economic_conservatism, sec_social_conservatism),
               names_to = "conservatism_type",
               names_prefix = "sec_",
               values_to = "conservatism_score") %>% 
  mutate(conservatism_type = recode(conservatism_type, `economic_conservatism` = "Economic conservatism",
                                    `social_conservatism` = "Social conservatism")) %>%
  pivot_longer(cols = agreeableness:emotional_stability,
               names_to = "personality_trait",
               values_to = "personality_score") %>% 
  mutate(personality_trait = recode(personality_trait, `agreeableness` = "Agreeableness",
                                    `conscientiousness` = "Conscientiousness",
                                    `openness_to_experiences` = "Openness to experiences",
                                    `extraversion` = "Extraversion",
                                    `emotional_stability` = "Emotional stability"))
```

```{r table1, echo = FALSE, results = "asis"}
tidy_table <- tidy %>% 
  mutate(age = case_when(       
      age <= 29 & age >= 20 ~ "[20-29]",
      age <= 39 & age >= 30 ~ "[30-39]",
      age <= 49 & age >= 40 ~ "[40-49]",
      age <= 59 & age >= 50 ~ "[50-59]",
      TRUE ~ "[60-68]"))
tidy_table <- tidy_table %>% 
  group_by(personality_trait, age) %>% 
  summarize(mean_personality_score = mean(personality_score)) %>% 
  pivot_wider(names_from = age, values_from = mean_personality_score) %>% 
  rename("Personality Trait" = personality_trait)

apa_table(tidy_table, 
            caption = "Mean Personality Trait Scores Across the Lifespan",
          col_spanners = list(`Age Group` = c(2, 6)),
          note = "Personality trait scores reported on a 1-7 scale."
)

#This table is helpful and conveys a lot of information in a very clear format. 
```

```{r figure1, fig.cap= "Generally, older adults rated themselves higher on each of the big five personality traits than young adults. This sample was unexpectedly introverted."}

ggplot(tidy, aes(x = age, y = personality_score, color = personality_trait)) +
  geom_smooth(se=FALSE, method = "lm")+
  labs(title = "Personality traits across the lifespan", subtitle = "Big 5 traits self-reported by adults aged 20-68",
       x= "Age", y = "How strongly participants identify with the personality trait", 
       color = "Personality Trait")+
  theme_minimal()+
  scale_color_viridis_d()

```

```{r figure2, fig.cap= "Generally, older adults reported stronger socially and fiscally conservative beliefs compared to young adults. This increase in conservative beliefs as we age may be more extreme for socially conservative beliefs than for fiscally conservate beliefs."}

ggplot(tidy, aes(x = age, y = conservatism_score, color = conservatism_type)) +
  geom_smooth(se=FALSE, method = "lm")+
  scale_color_viridis_d() +
  labs(title = "Conservatism across the lifespan (adults aged 20-68)", subtitle = "Difference in Economic and Social Conservatism by Gender",
       x= "Age", y = "Conservatism Score (higher = more conservative)",
       color = "Conservatism Type")+
  theme_minimal()+
  facet_wrap(~gender)

#You could use this same plot and facet wrap by ethnicity instead of gender. 
```

```{r messy_data}
messy <- tidy %>% 
  pivot_wider(
    names_from = conservatism_type, 
    values_from = conservatism_score) %>%
   pivot_wider(
     names_from = personality_trait,
     values_from = personality_score)

#we can use this messy data to run lm models
```

# Methods

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

# Results

# Discussion

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup