### 2019-12-10 scripts/manuscript.Rmd
edits to method results

---
title             : "Personality and political beliefs across the lifespan"
shorttitle        : "PERSONALITY AND POLITICAL BELIEFS"
author: 
  - name          : "Sarah Dimakis"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "University of Oregon, Eugene, OR"
    email         : "sdimakis@uoregon.edu"
  - name          : "Meghan Siritzky"
    affiliation   : "1"
  - name          : "Jamie Yellowtail"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "University of Oregon"
authornote: |
  This project was completed as part of the EDLD Introduction to Data Science class at the University of Oregon.
abstract: |
  Here is where we will write our abstract.

keywords          : "keywords"
wordcount         : "X"
bibliography      : ["r-references.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library(papaja)
library(rio)
library(here)
library(tidyverse)
library(psych)
library(knitr)
library(MBESS)
library(viridis)

data <- import(here("data", "EPB.sav"), setclass = "tbl_df")
```

```{r analysis-preferences}
knitr::opts_chunk$set(echo = FALSE)

```

```{r filter}
#Removing particpants from the sample

#filter out preview mode (not real participants) and those who did not complete the survey
data <- data %>% 
  filter(DistributionChannel != "preview" & Progress == 100)

#filter out those who did not pass the attention checks
data <- data %>% 
  filter(Q6_8 == 6 & Q5_14 == 3)

#filter out those who took the survey too quickly (the time cut offs were predetermined and based on which condition the participant was randomly assigned to)
data <- data %>% 
  pivot_longer(cols = Video_DO_V4:Video_DO_V1, 
               names_to = "Video_DO", 
               values_to = "Value",
               names_prefix = "Video_DO_",
               values_drop_na = TRUE) %>% 
  filter(Video_DO == 'V1' & Duration__in_seconds_ >316 | 
           Video_DO == 'V2' & Duration__in_seconds_ > 340 | 
           Video_DO == 'V3' & Duration__in_seconds_ > 398 |
           Video_DO == 'V4' & Duration__in_seconds_ > 666 | 
           Video_DO == 'V5' & Duration__in_seconds_ > 565 | 
           Video_DO == 'V6' & Duration__in_seconds_ > 538) %>% 
  filter(Duration__in_seconds_ < 5000) #this person spent too long taking the survey
```

```{r aggregrateSEC}

#Aggregating the social and economic conservationism scale (SECs) using scoreItems from the psych package and mutate from dplyr
items <- c("Q3_1", "Q3_2","Q3_3","Q3_4","Q3_5","Q3_6","Q3_7","Q3_8","Q3_9","Q3_10","Q3_11","Q3_12")

#econonmic conservatism subscale 
scaleKey <- c(0, 1, 0, 0, -1, 1, 0, 0, 1, 1, 0, 0)
economic <- scoreItems(keys = scaleKey, items = data[items], 
  totals = FALSE, missing = FALSE, min = 0, max = 100)
data <- data %>% 
  mutate(SEC_economic_conservatism = economic$scores[,1])

#social conservatism subscale
scaleKey <- c(-1, 0, 1, 1, 0, 0, 1, 1, 0, 0, 1, 1)
social <- scoreItems(keys = scaleKey, items = data[items], 
  totals = FALSE, missing = FALSE, min = 0, max = 100)
data <- data %>% 
  mutate(SEC_social_conservatism = social$scores[,1])
```

```{r aggregateTIPI}

#Aggregating big 5 traits (TIPI) using scoreItems from the psych package and mutate from dplyr
items <- c("Q4_1", "Q4_2", "Q4_3", "Q4_4", "Q4_5", "Q4_6", "Q4_7", "Q4_8", "Q4_9", "Q4_10")

#Extraversion
scaleKey <- c(1, 0, 0, 0, 0, -1, 0, 0, 0, 0)
extraversion <- scoreItems(keys = scaleKey, items = data[items], 
  totals = FALSE, missing = FALSE, min = 1, max = 7)
data <- data %>% 
  mutate(Extraversion = extraversion$scores[,1])

#Agreeableness
scaleKey <- c(0, -1, 0, 0, 0, 0, 1, 0, 0, 0)
agree <- scoreItems(keys = scaleKey, items = data[items], 
  totals = FALSE, missing = FALSE, min = 1, max = 7)
data <- data %>% 
  mutate(Agreeableness = agree$scores[,1])

#Conscientiousness
scaleKey <- c(0, 0, 1, 0, 0, 0, 0, -1, 0, 0)
consc <- scoreItems(keys = scaleKey, items = data[items], 
  totals = FALSE, missing = FALSE, min = 1, max = 7)
data <- data %>% mutate(Conscientiousness = consc$scores[,1])

#Emotional stability
scaleKey <- c(0, 0, 0, -1, 0, 0, 0, 0, 1, 0)
emo <- scoreItems(keys = scaleKey, items = data[items], 
  totals = FALSE, missing = FALSE, min = 1, max = 7)
data <- data %>% 
  mutate(Emotional_stability = emo$scores[,1])

#Openness to experiences
scaleKey <- c(0, 0, 0, 0, 5, 0, 0, 0, 0, -1)
open <- scoreItems(keys = scaleKey, items = data[items], 
  totals = FALSE, missing = FALSE, min = 1, max = 7)
data <- data %>% 
  mutate(Openness_to_experiences = open$scores[,1])
```

```{r select}
#selecting the variables that we want to use, cleaning, and characterizing dataset
data <- data %>% 
  select(ResponseId, Q7, Q8, Q9, Q10, Q11, Duration__in_seconds_,
         SEC_economic_conservatism, SEC_social_conservatism,
         Agreeableness, Conscientiousness, Openness_to_experiences, Extraversion, Emotional_stability) %>% 
  rename(Gender = Q7, 
         Age = Q8, 
         Ethnicity = Q9, 
         Political_orientation = Q10,
         Political_affiliation = Q11,
         id = ResponseId) %>% 
  janitor::clean_names() %>% 
  characterize() %>% 
  mutate(age = as.numeric(age))
```

```{r tidying}
#last step is to tidy our data
tidy <- data %>% 
  pivot_longer(cols = c(sec_economic_conservatism, sec_social_conservatism),
               names_to = "conservatism_type",
               names_prefix = "sec_",
               values_to = "conservatism_score") %>% 
  mutate(conservatism_type = recode(conservatism_type, `economic_conservatism` = "Economic conservatism",
                                    `social_conservatism` = "Social conservatism")) %>%
  pivot_longer(cols = agreeableness:emotional_stability,
               names_to = "personality_trait",
               values_to = "personality_score") %>% 
  mutate(personality_trait = 
           recode(personality_trait,
                  `agreeableness` = "Agreeableness",
                  `conscientiousness` = "Conscientiousness",
                  `openness_to_experiences` = "Openness to experiences",
                  `extraversion` = "Extraversion",
                  `emotional_stability` = "Emotional stability"))
```

```{r table1, echo = FALSE, results = "asis"}
#this table is of the mean personality trait scores by social conservatism scores
tidy_table_sc <- tidy %>% 
  filter(conservatism_type == 'Social conservatism') %>%
  mutate(conservatism_score = case_when(       
      conservatism_score <= 19 & conservatism_score >= 0 ~ "[0-19]",
      conservatism_score <= 39 & conservatism_score >= 20 ~ "[20-39]",
      conservatism_score <= 59 & conservatism_score >= 40 ~ "[40-59]",
      conservatism_score <= 79 & conservatism_score >= 60 ~ "[60-79]",
      TRUE ~ "[80-100]"))
tidy_table_sc <- tidy_table_sc %>%
  group_by(personality_trait, conservatism_score) %>% 
  summarize(mean_personality_score = mean(personality_score)) %>% 
  pivot_wider(names_from = conservatism_score, values_from = mean_personality_score)

apa_table(tidy_table_sc, 
            caption = "Mean Personality Trait Scores by Social Conservatism Score",
          col_spanners = list(`Social Conservatism Score` = c(2, 6)),
          note = "Personality trait scores reported on a 1-7 scale.")
```

```{r table2, echo = FALSE, results = "asis"}
tidy_table_ec <- tidy %>% 
  filter(conservatism_type == 'Economic conservatism') %>% 
  mutate(conservatism_score = case_when(       
      conservatism_score <= 19 & conservatism_score >= 0 ~ "[0-19]",
      conservatism_score <= 39 & conservatism_score >= 20 ~ "[20-39]",
      conservatism_score <= 59 & conservatism_score >= 40 ~ "[40-59]",
      conservatism_score <= 79 & conservatism_score >= 60 ~ "[60-79]",
      TRUE ~ "[80-100]"))
tidy_table_ec <- tidy_table_ec %>% 
  group_by(personality_trait, conservatism_score) %>% 
  summarize(mean_personality_score = mean(personality_score)) %>% 
  pivot_wider(names_from = conservatism_score, values_from = mean_personality_score)

apa_table(tidy_table_ec, 
            caption = "Mean Personality Trait Scores by Economic Conservatism Score",
          col_spanners = list(`Economic Conservatism Score` = c(2, 6)),
          note = "Personality trait scores reported on a 1-7 scale.")
```

```{r linearmodels_conservatism_ocean, include = FALSE}
untidy <- tidy %>% 
  pivot_wider(names_from = conservatism_type, values_from = conservatism_score) %>% 
  pivot_wider(names_from = personality_trait, values_from = personality_score) %>% 
  rename(social_conservatism = 'Social conservatism',
         economic_conservatism = 'Economic conservatism',
         openness_to_experiences = 'Openness to experiences',
         emotional_stability = 'Emotional stability')

sc_ocean <- lm(social_conservatism ~ openness_to_experiences + Conscientiousness + Extraversion + Agreeableness + emotional_stability,
              data = untidy)
summary(sc_ocean)
plot(scale(residuals(sc_ocean)) ~ scale(fitted(sc_ocean)), xlab="Social Conservatism Scores", ylab = "Scaled Residuals")

ec_ocean <- lm(economic_conservatism ~ openness_to_experiences + Conscientiousness + Extraversion + Agreeableness + emotional_stability,
              data = untidy)
summary(ec_ocean)
plot(scale(residuals(ec_ocean)) ~ scale(fitted(ec_ocean)), xlab="Economic Conservatism Scores", ylab = "Scaled Residuals")
```

```{r linearmodels_conservatism_ocean_tables, echo = FALSE, results = "asis"}
apa_sc_ocean <- apa_print(sc_ocean)
apa_sc_ocean$table %>% 
  apa_table(caption = "Regression Table Predicting Social Conservatism From Big-Five Personality Traits.", note = "Residual standard error: 25.13 on 125 degrees of freedom.
Multiple R-squared: 0.168, Adjusted R-squared: 0.135. F(5, 125): 5.051, p-value: 0.0003.", escape = FALSE)

apa_ec_ocean <- apa_print(ec_ocean)
apa_ec_ocean$table %>% 
  apa_table(caption = "Regression Table Predicting Economic Conservatism From Big-Five Personality Traits.", note = "Residual standard error: 21.3 on 125 degrees of freedom. Multiple R-squared: 0.081, Adjusted R-squared: 0.044. F(5, 125): 2.207, p-value: 0.058.", escape = FALSE)
```

# Methods

```{r, include = FALSE}
#Age statistics
mean_age <- mean(as.numeric(data$age)) %>% round(2)
sd_age <- sd(as.numeric(data$age)) %>% round(2)
r1_age <- range(as.numeric(data$age))[1]
r2_age <- range(as.numeric(data$age))[2]

#Number of participants
n <- nrow(data)

#Ethnicity
#I'm creating a function that will spit out the percentage of people with a given ethnicity, including a % sign
percent_ethn <- function(pos){
  paste(((table(data$ethnicity)[[pos]]/n) *100) %>% round(1), "%", sep = "")}
black <-percent_ethn(1)
asian <- percent_ethn(2)
white <- percent_ethn(3)
hispanic <- percent_ethn(4)
native <- percent_ethn(5)
other <- percent_ethn(6)

#Gender
percent_gen <- function(pos){
  paste(((table(data$gender)[[pos]]/n) *100) %>% round(1), "%", sep = "")}
female <- percent_gen(1)
male <- percent_gen(2)

#politicial affiliation
percent_pa <- function(pos){
  paste(((table(data$political_affiliation)[[pos]]/n) *100) %>% round(1), "%", sep = "")}
democrat <- percent_pa(1)
republican <- percent_pa(6)
no_party <- percent_pa(4)
other_n <- table(data$political_affiliation)[[2]]+
  table(data$political_affiliation)[[3]]+
  table(data$political_affiliation)[[5]]
other <- paste(((other_n/n) *100) %>% round(1), "%", sep = "")

#Duration
data <- data %>% 
  mutate(duration_in_seconds = as.numeric(duration_in_seconds))
duration_m <- mean(data$duration_in_seconds)/60
duration_sd <- sd(data$duration_in_seconds)/60
```
## Participants

Two-hundred U.S. adults were recruited via Amazon Mechanical Turk to participate in a study listed as "Psychology and society survey." Participants were removed for completing the survey quicker than predetermined time cutoffs or failing to pass attention checks. Of the remaining `r n` participants, who were `r r1_age` to `r r2_age` years old (*M* = `r mean_age`, *SD* = `r sd_age`), `r female` identified as female and `r male` identified as male, while `r white` identified as White of Caucasian, `r black` as African American or Black, `r hispanic` as Hispanic or Latinx, `r asian` as Asian or Asian American, `r native` as Native American, and `r other` as "Other." The sample leaned liberal, with `r democrat` affiliated with the Democratic Party, `r republican` affiliated with the Republican Party, `r no_party` affiliated with no party, and `r other` affiliated with the Green Party, Libertarian Party, or another party. The participants were compensated $1.50 for completing the survey, which took on average `r duration_m` minutes (*SD* = `r duration_sd`).

## Material

**Personality.** To assess personality traits (openness to experiences, conscientiousness, extraversion, agreeableness, and emotional stability), respondents completed the Ten Item Personality Inventory [@gosling2003very]. Participants were asked to indicate the extent to which they agreed or disagreed with statements describing themselves (e.g. “Extraverted, enthusiastic”; “Critical, quarrelsome”). Respondents rated each statement on a 7-point likert scale (1 = Strongly Disagree to 7 = Strongly Agree). Responses were coded such that higher values reflected greater identification with the personality trait.

**Social and Economic Conservatism.** To assess participant levels of conservatism, respondents completed the 12-item Social and Economic Conservatism Scale [@everett201312]. Participants were asked to indicate the extent to which they feel positively or negatively about seven social issues and five economic issues (eg. Abortion; Fiscal Responsibility). Respondents rated each issue on a 100-point scale (0 indicates greater negativity and 100 indicates greater positivity). Responses were coded such that higher values reflected greater levels of conservatism.

**Additional Measures.** Incidental affect was induced via video clips [@schaefer2010]. Clips from _The Shining_ and _The Blair Witch Project_ were used to induce fear, clips from _Seven_ and _Schindler’s List_ were used to induce anger, and clips from _Blue_ were used to induce no emotion. Participants also read a speech on the prevalence of homelessness in the United States and a corresponding policy varying in complexity and affiliation.

Participants completed additional measures, including an adjusted PANAS Affect Scale assessing current emotional state [@watson1988development], the 18-Point Need for Cognition Scale assessing need for cognition [@cacioppo1984efficient], the 10-Point Emotion Regulation Questionnaire assessing the use of two emotional regulation strategies, cognitive reappraisal and expressive supression [@gross2003individual], and demographic information, including gender, ethnicity, age, and political orientation and affiliation.

## Procedure

Participants accessed the survey through the Amazon Mechanical Turk website. Once informed consent was confirmed, participants were randomly assigned to watch a video intended to induce primarily fear, anger, or no emotion. After watching the video, each participant was presented with information describing the prevalance of homelessness in the United States. Participants were then randomly assigned to read a policy statement addressing homelessness that was either simple or complex, and liberal or conservative, and were then asked to answer questions measuring their beliefs about the policy. Additional measures followed measuring social and economic conservatism, personality traits, current affective state, neeed for cognition, and emotional regulation. Participants concluded the study by providing demographic information.

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

# Results

In order to explore relation between conservatism and personality, we created a table of the Big Five personality trait (openness to experiences, conscientiousness, extraversion, agreeableness, and emotional stability) scores across different levels of social (see Table 1) and economic (see Table 2) conservatism.

To see whether there were significant relations between social conservatism and the Big Five traits, we then ran a multiple regression predicting social conservatism from the Big Five personality traits. We found that together the personality traits explain about 13.5% of the variance in social conservatism scores, *F*(5, 125) = 5.05, *p* < .001, with an adjusted $R^2$ of .14 (see Table 3). We found that agreeablessness, $\beta$ = 3.93, *t*(125) = 2.40, *p* = .018, and emotional stability, $\beta$ = 4.30, *t*(125) = 2.37, *p* = .020, significantly predict social conservatism, such that, controlling for the other personality traits, those who scored higher on agreeableness and emotional stability measures tended to also score higher on the social conservatism measure. Additionally, openness to experiences margionally predicted social conservatism, $\beta$ = -3.73, *t*(125) = -1.96, *p* = .052, such that, controlling for other personality traits, those who scored higher on conservatism tended to score lower on the openness to experience measure.

To see whether there were significant relations between economic conservatism and the Big Five traits, we then ran a multiple regression predicting economic conservatism from the Big Five personality traits. The personality traits taken together did not significantly explain variance in economic conservatism scores, *F*(5, 125) = 2.21, *p* = .058 (see Table 4). However, openness to experience was a significant predictor of economic conservatism, $\beta$ = -3.79, *t*(125) = -2.35, *p* = .021, such that, controlling for other personality traits, those who scored high on economic conservatism subscale tended to score lower on the openness to experiences measure.

To better understand the relation between social and economic conservatism, we examined the correlation between social and economic conservatism scores, and found it to be *r* = .65. Neither social nor economic extraversion were strongly correlated with any of the Big Five personality traits (see Figure 1).

# Discussion

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>

```{r correlations, echo = FALSE, results = "asis"}
set.seed=180
cor_conservatism <- corr.test(untidy$economic_conservatism, untidy$social_conservatism)
cor_conservatism

untidy_cor <- untidy %>% 
  select(-c(1:7)) %>% 
  rename('Social conservatism' = 'social_conservatism',
         'Economic conservatism' = 'economic_conservatism',
         'Openness to experiences' = 'openness_to_experiences',
         'Emotional stability' = 'emotional_stability')
cor_conservatism_ocean <- round(cor(untidy_cor),2)
upper_s<-cor_conservatism_ocean
upper_s[upper.tri(cor_conservatism_ocean)]<-""
upper_s<-as.data.frame(upper_s) %>%
  apa_table(
    caption = "Correlation Matrix for Conservatism and Big Five Personality Traits",
  landscape = TRUE,
  font_size = "tiny")

```

```{r figure1, fig.cap="A correlation heatmap demonstrating the correlations between the Big Five personality traits and social and economic conservatism."}
heatmap(cor_conservatism_ocean,
        Rowv = NA,
        Colv = NA,
        main = "Correlation Heatmap",
        margins = c(15,15),
        col = viridis(100, direction = -1))
```

```{r figure2, fig.cap= "We found a significant positive relation between emotional stability and social conservatism, a positive relation between agreeableness and social conservatism, and a negative relation between openness to experiences and economic conservatism.", fig.height = 10, fig.width = 8}
ggplot(tidy, aes(x = personality_score, y = conservatism_score, color = conservatism_type)) +
  geom_smooth(se=FALSE, method = "lm")+
  labs(title = "Conservatism by Personality Traits", subtitle = "Big 5 traits self-reported by adults aged 20-68",
       x= "How strongly participants identify with the personality trait", y = "Conservatism score", 
       color = "Conservatism type")+
  theme_minimal()+
  scale_color_viridis_d() +
  facet_wrap(~personality_trait, ncol = 1)
```

```

\endgroup