### 2019-12-10 scripts/manuscript.tex

\documentclass[man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Personality and political beliefs across the lifespan},
            pdfauthor={Sarah Dimakis, Meghan Siritzky, \& Jamie Yellowtail},
            pdfkeywords={keywords},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{Personality and political beliefs across the lifespan}
    \author{Sarah Dimakis\textsuperscript{1}, Meghan Siritzky\textsuperscript{1}, \&
Jamie Yellowtail\textsuperscript{1}}
    \date{}

\shorttitle{PERSONALITY AND POLITICAL BELIEFS}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} University of Oregon}
\keywords{keywords\newline\indent Word count: X}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother

\authornote{This project was completed as part of the EDLD
Introduction to Data Science class at the University of Oregon.

Correspondence concerning this article should be addressed to Sarah
Dimakis, University of Oregon, Eugene, OR. E-mail:
\href{mailto:sdimakis@uoregon.edu}{\nolinkurl{sdimakis@uoregon.edu}}}

\abstract{
Here is where we will write our abstract.

}

\begin{document}
\maketitle

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:table1}Mean Personality Trait Scores by Social Conservatism Score}
\begin{tabular}{llllll}
\toprule
 & \multicolumn{5}{c}{Social Conservatism Score} \\
\cmidrule(r){2-6}
personality\_trait & \multicolumn{1}{c}{[0-19]} & \multicolumn{1}{c}{[20-39]} & \multicolumn{1}{c}{[40-59]} & \multicolumn{1}{c}{[60-79]} & \multicolumn{1}{c}{[80-100]}\\
\midrule
Agreeableness & 4.67 & 5.20 & 4.88 & 5.76 & 5.90\\
Conscientiousness & 5.37 & 5.44 & 5.32 & 5.84 & 6.03\\
Emotional stability & 4.07 & 4.38 & 4.82 & 5.50 & 5.57\\
Extraversion & 2.73 & 3.16 & 3.57 & 3.90 & 3.72\\
Openness to experiences & 4.93 & 5.40 & 5.43 & 5.25 & 5.12\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} Personality trait scores reported on a 1-7 scale.}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:table2}Mean Personality Trait Scores by Economic Conservatism Score}
\begin{tabular}{llllll}
\toprule
 & \multicolumn{5}{c}{Economic Conservatism Score} \\
\cmidrule(r){2-6}
personality\_trait & \multicolumn{1}{c}{[0-19]} & \multicolumn{1}{c}{[20-39]} & \multicolumn{1}{c}{[40-59]} & \multicolumn{1}{c}{[60-79]} & \multicolumn{1}{c}{[80-100]}\\
\midrule
Agreeableness & 6.14 & 5.21 & 5.29 & 5.40 & 5.43\\
Conscientiousness & 5.50 & 5.31 & 5.60 & 5.77 & 6.00\\
Emotional stability & 4.29 & 4.67 & 5.09 & 4.85 & 5.62\\
Extraversion & 2.79 & 3.67 & 3.61 & 3.42 & 3.50\\
Openness to experiences & 6.43 & 5.44 & 5.12 & 5.05 & 5.21\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} Personality trait scores reported on a 1-7 scale.}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:linearmodels_conservatism_ocean_tables}Regression Table Predicting Social Conservatism From Big-Five Personality Traits.}
\begin{tabular}{lllll}
\toprule
Predictor & \multicolumn{1}{c}{$b$} & \multicolumn{1}{c}{95\% CI} & \multicolumn{1}{c}{$t(125)$} & \multicolumn{1}{c}{$p$}\\
\midrule
Intercept & 25.82 & $[0.07$, $51.58]$ & 1.98 & .049\\
Openness to experiences & -3.73 & $[-7.50$, $0.04]$ & -1.96 & .052\\
Conscientiousness & 0.07 & $[-3.98$, $4.11]$ & 0.03 & .974\\
Extraversion & 1.56 & $[-1.42$, $4.55]$ & 1.04 & .302\\
Agreeableness & 3.93 & $[0.69$, $7.17]$ & 2.40 & .018\\
Emotional stability & 4.30 & $[0.70$, $7.91]$ & 2.37 & .020\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} Residual standard error: 25.13 on 125 degrees of freedom.
Multiple R-squared: 0.168, Adjusted R-squared: 0.135. F(5, 125): 5.051, p-value: 0.0003.}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:linearmodels_conservatism_ocean_tables}Regression Table Predicting Economic Conservatism From Big-Five Personality Traits.}
\begin{tabular}{lllll}
\toprule
Predictor & \multicolumn{1}{c}{$b$} & \multicolumn{1}{c}{95\% CI} & \multicolumn{1}{c}{$t(125)$} & \multicolumn{1}{c}{$p$}\\
\midrule
Intercept & 56.54 & $[34.71$, $78.37]$ & 5.13 & < .001\\
Openness to experiences & -3.79 & $[-6.98$, $-0.59]$ & -2.35 & .021\\
Conscientiousness & 1.78 & $[-1.65$, $5.20]$ & 1.03 & .307\\
Extraversion & 0.61 & $[-1.92$, $3.14]$ & 0.48 & .632\\
Agreeableness & -1.13 & $[-3.88$, $1.61]$ & -0.82 & .415\\
Emotional stability & 2.34 & $[-0.71$, $5.39]$ & 1.52 & .131\\
\bottomrule
\addlinespace
\end{tabular}
\begin{tablenotes}[para]
\normalsize{\textit{Note.} Residual standard error: 21.3 on 125 degrees of freedom. Multiple R-squared: 0.081, Adjusted R-squared: 0.044. F(5, 125): 2.207, p-value: 0.058.}
\end{tablenotes}
\end{threeparttable}
\end{center}
\end{table}

\section{Methods}\label{methods}

\subsection{Participants}\label{participants}

Two-hundred U.S. adults were recruited via Amazon Mechanical Turk to
participate in a study listed as \enquote{Psychology and society
survey.} Participants were removed for completing the survey quicker
than predetermined time cutoffs or failing to pass attention checks. Of
the remaining 131 participants, who were 20 to 68 years old (\emph{M} =
40.55, \emph{SD} = 11.98), 54.2\% identified as female and 45.8\%
identified as male, while 77.1\% identified as White of Caucasian, 7.6\%
as African American or Black, 6.1\% as Hispanic or Latinx, 5.3\% as
Asian or Asian American, 0.8\% as Native American, and 9.2\% as
\enquote{Other.} The sample leaned liberal, with 45\% affiliated with
the Democratic Party, 23.7\% affiliated with the Republican Party,
22.1\% affiliated with no party, and 9.2\% affiliated with the Green
Party, Libertarian Party, or another party. The participants were
compensated \$1.50 for completing the survey, which took on average
10.43 minutes (\emph{SD} = 3.34).

\subsection{Material}\label{material}

\textbf{Personality.} To assess personality traits (openness to
experiences, conscientiousness, extraversion, agreeableness, and
emotional stability), respondents completed the Ten Item Personality
Inventory (Gosling, Rentfrow, \& Swann Jr, 2003). Participants were
asked to indicate the extent to which they agreed or disagreed with
statements describing themselves (e.g. \enquote{Extraverted,
enthusiastic}; \enquote{Critical, quarrelsome}). Respondents rated each
statement on a 7-point likert scale (1 = Strongly Disagree to 7 =
Strongly Agree). Responses were coded such that higher values reflected
greater identification with the personality trait.

\textbf{Social and Economic Conservatism.} To assess participant levels
of conservatism, respondents completed the 12-item Social and Economic
Conservatism Scale (Everett, 2013). Participants were asked to indicate
the extent to which they feel positively or negatively about seven
social issues and five economic issues (eg. Abortion; Fiscal
Responsibility). Respondents rated each issue on a 100-point scale (0
indicates greater negativity and 100 indicates greater positivity).
Responses were coded such that higher values reflected greater levels of
conservatism.

\textbf{Additional Measures.} Incidental affect was induced via video
clips (Schaefer, Nils, Sanchez, \& Philippot, 2010). Clips from
\emph{The Shining} and \emph{The Blair Witch Project} were used to
induce fear, clips from \emph{Seven} and \emph{Schindler's List} were
used to induce anger, and clips from \emph{Blue} were used to induce no
emotion. Participants also read a speech on the prevalence of
homelessness in the United States and a corresponding policy varying in
complexity and affiliation.

Participants completed additional measures, including an adjusted PANAS
Affect Scale assessing current emotional state (Watson, Clark, \&
Tellegen, 1988), the 18-Point Need for Cognition Scale assessing need
for cognition (Cacioppo, Petty, \& Feng Kao, 1984), the 10-Point Emotion
Regulation Questionnaire assessing the use of two emotional regulation
strategies, cognitive reappraisal and expressive supression (Gross \&
John, 2003), and demographic information, including gender, ethnicity,
age, and political orientation and affiliation.

\subsection{Procedure}\label{procedure}

Participants accessed the survey through the Amazon Mechanical Turk
website. Once informed consent was confirmed, participants were randomly
assigned to watch a video intended to induce primarily fear, anger, or
no emotion. After watching the video, each participant was presented
with information describing the prevalance of homelessness in the United
States. Participants were then randomly assigned to read a policy
statement addressing homelessness that was either simple or complex, and
liberal or conservative, and were then asked to answer questions
measuring their beliefs about the policy. Additional measures followed
measuring social and economic conservatism, personality traits, current
affective state, neeed for cognition, and emotional regulation.
Participants concluded the study by providing demographic information.

\subsection{Data analysis}\label{data-analysis}

We used R (Version 3.6.1; R Core Team, 2019) and the R-packages
\emph{\}MBESS} {[}@\}R-MBESS{]}, \emph{dplyr} (Version 0.8.3; Wickham,
François, Henry, \& Müller, 2019), \emph{forcats} (Version 0.4.0;
Wickham, 2019a), \emph{ggplot2} (Version 3.2.1; Wickham, 2016),
\emph{here} (Version 0.1; Müller, 2017), \emph{janitor} (Version 1.2.0;
Firke, 2019), \emph{knitr} (Version 1.25; Xie, 2015), \emph{papaja}
(Version 0.1.0.9842; Aust \& Barth, 2018), \emph{psych} (Version 1.8.12;
Revelle, 2018), \emph{purrr} (Version 0.3.3; Henry \& Wickham, 2019),
\emph{readr} (Version 1.3.1; Wickham, Hester, \& Francois, 2018),
\emph{rio} (Version 0.5.16; C.-h. Chan, Chan, Leeper, \& Becker, 2018),
\emph{stringr} (Version 1.4.0; Wickham, 2019b), \emph{tibble} (Version
2.1.3; Müller \& Wickham, 2019), \emph{tidyr} (Version 1.0.0; Wickham \&
Henry, 2019), \emph{tidyverse} (Version 1.2.1; Wickham, 2017),
\emph{viridis} (Version 0.5.1; Garnier, 2018a, 2018b), and
\emph{viridisLite} (Version 0.3.0; Garnier, 2018b) for all our analyses.

\section{Results}\label{results}

In order to explore relation between conservatism and personality, we
created a table of the Big Five personality trait (openness to
experiences, conscientiousness, extraversion, agreeableness, and
emotional stability) scores across different levels of social (see Table
1) and economic (see Table 2) conservatism.

To see whether there were significant relations between social
conservatism and the Big Five traits, we then ran a multiple regression
predicting social conservatism from the Big Five personality traits. We
found that together the personality traits explain about 13.5\% of the
variance in social conservatism scores, \emph{F}(5, 125) = 5.05,
\emph{p} \textless{} .001, with an adjusted \(R^2\) of .14 (see Table
3). We found that agreeablessness, \(\beta\) = 3.93, \emph{t}(125) =
2.40, \emph{p} = .018, and emotional stability, \(\beta\) = 4.30,
\emph{t}(125) = 2.37, \emph{p} = .020, significantly predict social
conservatism, such that, controlling for the other personality traits,
those who scored higher on agreeableness and emotional stability
measures tended to also score higher on the social conservatism measure.
Additionally, openness to experiences margionally predicted social
conservatism, \(\beta\) = -3.73, \emph{t}(125) = -1.96, \emph{p} = .052,
such that, controlling for other personality traits, those who scored
higher on conservatism tended to score lower on the openness to
experience measure.

To see whether there were significant relations between economic
conservatism and the Big Five traits, we then ran a multiple regression
predicting economic conservatism from the Big Five personality traits.
The personality traits taken together did not significantly explain
variance in economic conservatism scores, \emph{F}(5, 125) = 2.21,
\emph{p} = .058 (see Table 4). However, openness to experience was a
significant predictor of economic conservatism, \(\beta\) = -3.79,
\emph{t}(125) = -2.35, \emph{p} = .021, such that, controlling for other
personality traits, those who scored high on economic conservatism
subscale tended to score lower on the openness to experiences measure.

To better understand the relation between social and economic
conservatism, we examined the correlation between social and economic
conservatism scores, and found it to be \emph{r} = .65. Neither social
nor economic extraversion were strongly correlated with any of the Big
Five personality traits (see Figure 1).

\section{Discussion}\label{discussion}

\newpage

\section{References}\label{references}

\begin{verbatim}
## Warning in readLines(file): incomplete final line found on 'r-
## references.bib'
\end{verbatim}

\begingroup
\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\hypertarget{ref-R-papaja}{}
Aust, F., \& Barth, M. (2018). \emph{papaja: Create APA manuscripts with
R Markdown}. Retrieved from \url{https://github.com/crsh/papaja}

\hypertarget{ref-cacioppo1984efficient}{}
Cacioppo, J. T., Petty, R. E., \& Feng Kao, C. (1984). The efficient
assessment of need for cognition. \emph{Journal of Personality
Assessment}, \emph{48}(3), 306--307.

\hypertarget{ref-R-rio}{}
Chan, C.-h., Chan, G. C., Leeper, T. J., \& Becker, J. (2018).
\emph{Rio: A swiss-army knife for data file i/o}.

\hypertarget{ref-everett201312}{}
Everett, J. A. (2013). The 12 item social and economic conservatism
scale (secs). \emph{PloS One}, \emph{8}(12), e82131.

\hypertarget{ref-R-janitor}{}
Firke, S. (2019). \emph{Janitor: Simple tools for examining and cleaning
dirty data}. Retrieved from
\url{https://CRAN.R-project.org/package=janitor}

\hypertarget{ref-R-viridis}{}
Garnier, S. (2018a). \emph{Viridis: Default color maps from
'matplotlib'}. Retrieved from
\url{https://CRAN.R-project.org/package=viridis}

\hypertarget{ref-R-viridisLite}{}
Garnier, S. (2018b). \emph{ViridisLite: Default color maps from
'matplotlib' (lite version)}. Retrieved from
\url{https://CRAN.R-project.org/package=viridisLite}

\hypertarget{ref-gosling2003very}{}
Gosling, S. D., Rentfrow, P. J., \& Swann Jr, W. B. (2003). A very brief
measure of the big-five personality domains. \emph{Journal of Research
in Personality}, \emph{37}(6), 504--528.

\hypertarget{ref-gross2003individual}{}
Gross, J. J., \& John, O. P. (2003). Individual differences in two
emotion regulation processes: Implications for affect, relationships,
and well-being. \emph{Journal of Personality and Social Psychology},
\emph{85}(2), 348.

\hypertarget{ref-R-purrr}{}
Henry, L., \& Wickham, H. (2019). \emph{Purrr: Functional programming
tools}. Retrieved from \url{https://CRAN.R-project.org/package=purrr}

\hypertarget{ref-R-here}{}
Müller, K. (2017). \emph{Here: A simpler way to find your files}.
Retrieved from \url{https://CRAN.R-project.org/package=here}

\hypertarget{ref-R-tibble}{}
Müller, K., \& Wickham, H. (2019). \emph{Tibble: Simple data frames}.
Retrieved from \url{https://CRAN.R-project.org/package=tibble}

\hypertarget{ref-R-base}{}
R Core Team. (2019). \emph{R: A language and environment for statistical
computing}. Vienna, Austria: R Foundation for Statistical Computing.
Retrieved from \url{https://www.R-project.org/}

\hypertarget{ref-R-psych}{}
Revelle, W. (2018). \emph{Psych: Procedures for psychological,
psychometric, and personality research}. Evanston, Illinois:
Northwestern University. Retrieved from
\url{https://CRAN.R-project.org/package=psych}

\hypertarget{ref-schaefer2010}{}
Schaefer, A., Nils, F., Sanchez, X., \& Philippot, P. (2010). Assessing
the effectiveness of a large database of emotion-eliciting films: A new
tool for emotion researchers. \emph{Cognition and Emotion},
\emph{24}(7), 1153--1172.

\hypertarget{ref-watson1988development}{}
Watson, D., Clark, L. A., \& Tellegen, A. (1988). Development and
validation of brief measures of positive and negative affect: The panas
scales. \emph{Journal of Personality and Social Psychology},
\emph{54}(6), 1063.

\hypertarget{ref-R-ggplot2}{}
Wickham, H. (2016). \emph{Ggplot2: Elegant graphics for data analysis}.
Springer-Verlag New York. Retrieved from
\url{https://ggplot2.tidyverse.org}

\hypertarget{ref-R-tidyverse}{}
Wickham, H. (2017). \emph{Tidyverse: Easily install and load the
'tidyverse'}. Retrieved from
\url{https://CRAN.R-project.org/package=tidyverse}

\hypertarget{ref-R-forcats}{}
Wickham, H. (2019a). \emph{Forcats: Tools for working with categorical
variables (factors)}. Retrieved from
\url{https://CRAN.R-project.org/package=forcats}

\hypertarget{ref-R-stringr}{}
Wickham, H. (2019b). \emph{Stringr: Simple, consistent wrappers for
common string operations}. Retrieved from
\url{https://CRAN.R-project.org/package=stringr}

\hypertarget{ref-R-tidyr}{}
Wickham, H., \& Henry, L. (2019). \emph{Tidyr: Tidy messy data}.
Retrieved from \url{https://CRAN.R-project.org/package=tidyr}

\hypertarget{ref-R-dplyr}{}
Wickham, H., François, R., Henry, L., \& Müller, K. (2019). \emph{Dplyr:
A grammar of data manipulation}. Retrieved from
\url{https://CRAN.R-project.org/package=dplyr}

\hypertarget{ref-R-readr}{}
Wickham, H., Hester, J., \& Francois, R. (2018). \emph{Readr: Read
rectangular text data}. Retrieved from
\url{https://CRAN.R-project.org/package=readr}

\hypertarget{ref-R-knitr}{}
Xie, Y. (2015). \emph{Dynamic documents with R and knitr} (2nd ed.).
Boca Raton, Florida: Chapman; Hall/CRC. Retrieved from
\url{https://yihui.name/knitr/}

Call:corr.test(x =
untidy\(economic_conservatism, y = untidy\)social\_conservatism)
Correlation matrix {[}1{]} 0.65 Sample Size {[}1{]} 131 Probability
values adjusted for multiple tests. {[}1{]} 0

To see confidence intervals of the correlations, print with the
short=FALSE option

\begin{lltable}

\tiny{
\begin{longtable}{llllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
\caption{\label{tab:correlations}Correlation Matrix for Conservatism and Big Five Personality Traits}\\
\toprule
 & \multicolumn{1}{c}{Economic conservatism} & \multicolumn{1}{c}{Social conservatism} & \multicolumn{1}{c}{Agreeableness} & \multicolumn{1}{c}{Conscientiousness} & \multicolumn{1}{c}{Openness to experiences} & \multicolumn{1}{c}{Extraversion} & \multicolumn{1}{c}{Emotional stability}\\
\midrule
Economic conservatism & 1 &  &  &  &  &  & \\
Social conservatism & 0.65 & 1 &  &  &  &  & \\
Agreeableness & -0.02 & 0.29 & 1 &  &  &  & \\
Conscientiousness & 0.14 & 0.18 & 0.28 & 1 &  &  & \\
Openness to experiences & -0.15 & -0.02 & 0.2 & 0.18 & 1 &  & \\
Extraversion & 0.02 & 0.17 & 0.2 & 0.12 & 0.44 & 1 & \\
Emotional stability & 0.16 & 0.33 & 0.35 & 0.54 & 0.28 & 0.42 & 1\\
\bottomrule
\end{longtable}
}
\end{lltable}

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/figure1-1.pdf}
\caption{\label{fig:figure1}A correlation heatmap demonstrating the
correlations between the Big Five personality traits and social and
economic conservatism.}
\end{figure}

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/figure2-1.pdf}
\caption{\label{fig:figure2}We found a significant positive relation between
emotional stability and social conservatism, a positive relation between
agreeableness and social conservatism, and a negative relation between
openness to experiences and economic conservatism.}
\end{figure}

```

\endgroup

\end{document}